---
title: "Halloween Mini Project 103024"
author: "Janie Changberg (PID: A69037446)"
format: pdf
---

```{r}
candy_file <- "candy-data.csv"
candy = read.csv(candy_file,row.names=1)
head(candy)
```
```{r}
dim(candy)
nrow(candy)
sum(candy$fruity)
```

##Q1. How many different candy types are in this dataset?

> 85 different candies

##Q2. How many fruity candy types are in the dataset?

> 38 types of fruity candy

```{r}
candy["Twix", ]$winpercent
```

```{r}
candy["Reese's Peanut Butter cup",]$winpercent
candy["Kit Kat",]$winpercent
candy["Tootsie Roll Snack Bars",]$winpercent
```

##Q3. What is your favorite candy in the dataset and what is its `winpercent` value?

> Reese's Peanut Butter Cups, 84.18029.

##Q4. What is the `winpercent` value for “Kit Kat”?

> 76.768

##Q5. What is the `winpercent` value for “Tootsie Roll Snack Bars”? \>49.6535

Installed Skimr:

```{r}
library("skimr")
skim(candy)
```

##Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

>The three "percent" values (sugarpercent, pricepercent, winpercent) are continuous variables vs binary values (the other columns are 0 or 1) for the p0-p100 values.


##Q7. What do you think a zero and one represent for the candy$chocolate column?

>They likely represent a "yes/no" answer to whether that particular candy contains that element. 

##Q8. Plot a histogram of `winpercent` values
```{r}
library(ggplot2)

ggplot(candy, aes(x=winpercent))+
  geom_histogram(binwidth=5, color=("black"))
```

##Q9. Is the distribution of `winpercent` values symmetrical?

>No, the values are skewed to the right.

##Q10. Is the center of the distribution above or below 50%?

>The center of the distribution is below 50%.

##Q11. On average is chocolate candy higher or lower ranked than fruit candy?
```{r}
choc_mean <- mean(candy$winpercent[as.logical(candy$chocolate)], na.rm=TRUE)
fruit_mean <- mean(candy$winpercent[as.logical(candy$fruity)], na.rm=TRUE)

choc_mean
fruit_mean
```
>Chocolate candy is ranked higher than fruit candy (60.92% win vs 44.12% win)

##Q12. Is this difference statistically significant?

```{r}
ttest_chocvsfruit <- t.test(candy$winpercent[as.logical(candy$chocolate)], na.rm=TRUE,candy$winpercent[as.logical(candy$fruity)], na.rm=TRUE)

ttest_chocvsfruit

```
>Yes, this difference is statistically significant, p-val=2.871e-08
  
##Q13. what are the five least liked candy types in this set?
```{r}
inds <- order(candy$winpercent)
head(candy[inds,])
```
>Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, Jawbusters
    

##Q14. What are the top 5 all time favorite candy types out of this set?
```{r}
inds2 <- order(candy$winpercent)
tail(candy[inds,])
```
> Reese's pieces, Snickers, Kit Kat, Twix, Reese's Miniatures

##Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
q15plot<- ggplot(candy)+
  aes(winpercent, rownames(candy))+
  geom_col()
q15plot
```

##Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?
```{r}
q16plot <- ggplot(candy)+
  aes(winpercent, reorder(rownames(candy),winpercent))+
  geom_col()
q16plot
```
##Let's make it purdy
```{r}
my_cols=rep("orange", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "brown"
my_cols[as.logical(candy$bar)] = "blue"
my_cols[as.logical(candy$fruity)] = "darkgreen"

beautify <- ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
beautify
```
##Q17. What is the worst ranked chocolate candy?
>The worst ranked chocolate candy is sixlets (as an aside, I have never heard of these).

##Q18. What is the best ranked fruity candy?
>The best ranked fruity candy is starbursts.

##What about value for money? What is the the best candy for the least money? 
```{r}
library(ggrepel)


bestval <- ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
bestval
```

##Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

> Reese's Miniatures are around the 25th percentile in price, and is the second most popular candy.

##Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
my_cols=rep("orange", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "brown"
my_cols[as.logical(candy$bar)] = "blue"
my_cols[as.logical(candy$fruity)] = "darkgreen"

q20beautify <- ggplot(candy) + 
  aes(pricepercent, reorder(rownames(candy),pricepercent)) +
  geom_col(fill=my_cols) 
q20beautify
```

>Nik L Nip, Nestle Smarties, Ring Pop, Mr. Good bar, and Hershey's Special Dark. The least popular of these is Nik L Nip.

##Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

>see plot generated for Q20 for barplot based on pricepercent. Below is the called for improvement:

```{r}
lollipoppriceplot <- ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), xend = 0), col="gray40") +
  geom_point()
lollipoppriceplot
```

##Exploring the Correlation Structure
```{r}
library(corrplot)

corstruct <- cor(candy)
corrplot(corstruct)
```
##Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

> Chocolate and Fruity, bar and pluribus.

##Q23. Similarly, what two variables are most positively correlated?

> Chocolate & bar, and Chocolate & winpercent

##Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=18)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

##Make nice with GGPlot:
```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```
##Use GGRepel to make labels that don't overlap

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

##Still weird, use plotly
```{r}
library(plotly)
```

```{r}
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

##Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

> Fruity and Pluribus. Yes, fruity candy often is sold as multiples in a packet.