---
title: "JCW_110824_GenomeInformatics_Class12"
author: "Janie Chang-Weinberg (A69037446)"
format: gfm
---

Looking through MXL .csv

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

```{r}
tbl.mxl <- table(mxl$Genotype..forward.strand.)
tbl.mxl
```

> There are 9 instances of G\|G out of 64 genotypes. 14.06% are homozygous for the asthma-related SNP.

## Homework (Questions 13 and 14)

Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

Read txt file into .csv

```{r}
genos <- read.table("https://bioboot.github.io/bggn213_F24/class-material/rs8067378_ENSG00000172057.6.txt")

head(genos)
```

```{r}
summary(genos)
```

```{r}
genosplot <- boxplot(genos$exp ~ genos$geno)
```

```{r}
genosplot
```

> Sample sizes: A/A: 108 A/G: 233 G/G: 121

> Median expression levels: A/A: 31.24847 A/G: 25.06486 G/G: 20.07363

Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

> (see above for boxplot, I should have read ahead, oops) The overall expression levels of A/A is greater than G/G, indicating that the SNP negatively impacts expression of G/G vs A/A
