---
title: "JCW_110624_Class11-MachineLearning2"
author: "Janie Chang-Weinberg (A69037446)"
format: pdf
---

Today, before delving into structure prediction with AlphaFold, we will finish off our previous lab 10 "comparative structure analysis" section.

```{r}
library(bio3d)

id <- "1ake_A"

aa <- get.seq(id)
aa
```

```{r}
b <- blast.pdb(aa)
```

```{r}
attributes(b)
```

```{r}
head(b$hit.tbl)
```

```{r}
hits <- plot(b)
```

```{r}
attributes(hits)
```

Top hits that we like from our blast results:
```{r}
hits$pdb.id
```

```{r}
#Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

I have now found and downloaded all ADK structures in the PDB database but viewing them is difficult as they need to be aligned and superposed. 

I am going to install the BiocManager from CRAN, then I can use `BiocManager::instal()` to install any bioconductor package. 

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
pdbs
```

##Principal Component Analysis
```{r}
pc <- pca(pdbs)
plot(pc)
```
fig 4 here is the scree plot! "Eigenvalue" are how you order the PCs. Eigenvalue 1 is PC1, Eigenvalue 2 is PC2, etc.  

```{r}
plot(pc, pc.axes=c(1:2))
```

To examine in more detail what PC1 (or any PC) is capturing here, we can plot the loadings or make an animated file of moving along PC1

```{r}
mktrj(pc, pc=1, file="pc1.pdb")
```

