---
title: "JCW_Class8_BreastCancer-MiniProject"
author: "Janie Chang-Weinberg (A69037446)"
format: pdf

---

Load and save data under new name:

```{r}
# Save your input data file into your Project directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data, row.names=1)
```

```{r}
wisc.df
```

Remove the first column!
```{r}
# We can use -1 here to remove the first column
wisc.data <- wisc.df[,-1]
```

check to make sure first column is gone:
```{r}
wisc.data
```

Create a diagnosis vector for later:
```{r}
diagnosis <- wisc.df$diagnosis

#just to check:
diagnosis
```

#Q1. How many observations are in this dataset?
```{r}
dim(wisc.data)
```
> There are 569 observations.

#Q2. How many of the observations have a malignant diagnosis?
```{r}
mal <- grep("M", diagnosis)
length(mal)
```
> 212 observations have a malignant diagnosis.

#Q3. How many variables/features in the data are suffixed with _mean?
> 10

Section 2: Principal Component Analysis
```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)

```
```{r}
wisc.pr <- prcomp( ___ )
```

