---
---
title: "Lab 6"
author: "Janie Chang-Weinberg (PID: A69037446)"
format: pdf
editor: visual
---
---

```{r}
add <- function(x,y=1,z=0){
  x + y
}
```

```{r}
add(1,1)

```

```{r}
add(1,1,z=1)
```

Q1: Define a function "Generate_DNA()" that makes a random nucleotide sequence of any length

# A second Function

```{r}
#generate_dna <- function()
bases <- c("A","C","T","G")
sequence <- sample(bases, replace=TRUE, size=5)
```
^^the above are code snippets that can be stitched together for function building


```{r}
generate_dna <- function(length) {
  bases <- c("A","C","T","G")
  sequence <- sample(bases, size=length, 
                    replace=TRUE)
  return(sequence)
  
}
```

```{r}
generate_dna(10)
```
tadaa

```{r}
aas <- unique(bio3d::aa.table$aa1)[1:20]
aaseq <- sample(aas, size=5, replace= TRUE)
```


Now make a similar function to the "generate_dna" function, but with amino acids.

```{r}
generate_pep <- function(length){
  aas <- unique(bio3d::aa.table$aa1)[1:20]
  aaseq <- sample(aas,size=length,replace=TRUE)
  aaseq <- paste(aaseq, collapse="") #collapse command will remove spaces between returned vals
  return(aaseq)
}
```

Q2: Generate random protein sequences of length 6-13 AAs.

```{r}
generate_pep(6)
```
Generate random sequences of length 6-12 AAs
```{r}
answer <- sapply(6:12, generate_pep)
answer
```

```{r}
cat(paste(">id.", 6:12, "\n", answer, sep=""),sep="\n")
```
